---
title: "Road To The NBA Playoffs"
author: "Pranav Vogeti, Niko Klekas"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Introduction

[placeholder]

### The Focus Question

[placeholder]

### Motivation

[placeholder]


# Background

### About the Original Data

[placeholder]


**NOTE** This data will be made available to our Github ^[Link to Github: https://github.com/psvogeti/SportsStatsDiscovery/tree/data]


### Data Description

[placeholder]

### The Key Variables

* The key variables from this data set we will use for this excursion: 
  - `var_name` variable definition
  

### Assumptions and Unusuality

[placeholder] 



### Report Intentions

[placeholder]


# Analysis

### Data Cleanup

After cleaning, the first several rows of our data set look like this: 
```{r, echo = FALSE}
# import the data
raw_draft_data <- read.csv("partner-data/nba-lottery-draft-data.csv")

format_draft_data <- raw_draft_data %>%
  rename(
    team = Team, 
    season = Season,
    season_outcome = Season_Outcome, 
    lottery_status = Lottery_Status,
    lottery_pick_count = Lottery_Pick_Count
  )
#comment
```

### Data Summaries

[placeholder]

```{r, echo = FALSE}
## placeholder
```

### Data Analysis


#### [placeholder]

[placeholder]

```{r, echo = FALSE}
## placeholder
```

### Discussion

[placeholder]

### Overall Conclusions


#### Shortcomings and Limitations

[placeholder]

#### Alternate Approaches

[placeholder]

#### Potential Future Directions

[placeholder]

git
# References
