---
title: "Three_Pointers"
author: "Pranav Vogeti, Niko Klekas"
date: "2023-06-05"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # convenient wrapper library for our data parsing
library(readxl) # this is to read in our excel (.xlsx) data
library(kableExtra) # this is to format our outputted tables #hello
```



# Introduction

[Placeholder]

### The Focus Question

[Placeholder]

### Motivation

[Placeholder]


# Background

### About the Original Data

[Placeholder]

### Data Description

[Placeholder]

### The Key Variables

* The key variables from this data set we will use for this excursion: 
  - `var` defintion
 
  

### Assumptions and Unusuality

[Placeholder]



### Report Intentions

[Placeholder]



# Analysis

### Data Cleanup

[Placeholder]

```{r}

```


### Data Summaries

[Placeholder]


```{r}

```

### Data Analysis

```{r}

```


# Discussion

[Placeholder]

### Overall Conclusions

#### [Placeholder] Discussion

[Placeholder]

### Shortcomings and Limitations

[Placeholder]


### Alternate Approaches

[Placeholder]


### Potential Future Directions

[Placeholder]


# References
